Cryptanalysis of \MORUS

### 2018-05-10 morusAC_04_Mini_trails.tex

Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive steps: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice. Then $\alpha^t_i$ is used one final time to close the trail. One may naturally wonder if some components of this trail are in conflict. In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$. To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.

>>>>

Of course, the overall trail resulting from all of the previous combinations is quite complex, especially since $\gamma^t_i$, $\delta^t_i$, and $\varepsilon^t_i$ each require two copies of the preceding trail fragment in consecutive steps: that is, $\varepsilon^t_i$ requires two approximations of $S_2$, which requires using $\delta^t_i$ twice; and $\delta^t_i$ in turn requires using $\gamma^t_i$ twice, which itself requires using $\alpha^t_i$ and $\beta^t_i$ twice. Then $\alpha^t_i$ is used one final time to close the trail.
The full construction with the exact bit indices for \MiniMORUS[640] and
\MiniMORUS[1280] is illustrated in \Cref{fig:twoapproximations}, where the left
and right half each show half of the full trail.
One may naturally wonder if some components of this trail are in conflict. In particular, products of bits from registers $S_2$ and $S_3$ are approximated multiple times, by $\alpha^t_i$, $\beta^t_i$ and $\gamma^t_i$. To address this concern, and ensure that all approximations along the trail are in fact compatible, we now compute the full trail equation explicitly.

\begin{figure}
  \newcommand{\M}[2]{1.65,-1.5*#1-.25-.25*#2}
  \newcommand{\C}[1]{-.1,-.25-1.5*#1}
  \renewcommand{\S}[2]{.25+.25*#2,-1.5*#1}
  \centering
  \resizebox{!}{0.95\textheight}{
  \begin{tikzpicture}[xscale=3, yscale=1.5, every node/.style={font=\scriptsize,inner sep=2pt}]
    \begin{scope}
      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S_{\r}$}; }
      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};

      \draw[alpha] (\C{0}) node[left] {27} -| (\S{1}{0}) node[above right] {0}
                   (\M{0}{0}) node[right] {$\alpha_{27}$};

      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
                   (\M{1}{0}) node[right] {$\beta_0$};
      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {8} -| (\S{2}{0}) node[above right] {\!\!13}
                   (\M{1}{1}) node[right] {$\alpha_{8,26}$};
      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {26} -| ([xshift=-1pt]\S{2}{0}) node[above left] {31\!};
      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {31\!\!}
                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
                   (\M{1}{2}) node[right] {$\gamma_0$};
      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {13}
                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
                   (\M{1}{3}) node[right] {$\delta_0$};

      \draw[beta]  (\C{2}) node[above left] {31} -| ([xshift=-1pt]\S{2}{0}) node[below left] {31\!}
                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {31\!\!}
                   (\M{2}{0}) node[right] {$\beta_{13,31}$};
      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {13} -| (\S{2}{0}) node[below right] {\!\!13}
                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,13};
      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {7} -| (\S{3}{0}) node[above right] {12}
                   (\M{2}{1}) node[right] {$\alpha_7$};
      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {13} -- ([xshift=2pt]\S{3}{1}) node[above right] {12}
                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {13}
                   (\M{2}{2}) node[right] {$\gamma_{13}$};

      \draw[beta]  (\C{3}) node[left] {12} -| (\S{3}{0}) node[below right] {12}
                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {12}
                   (\M{3}{0}) node[right] {$\beta_{12}$};

      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
    \end{scope}

    \begin{scope}[xshift=2.5cm]
      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S_{\r}$}; }
      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};

      \draw[alpha] (\C{1}) node[left] {2} -| (\S{2}{0}) node[above right] {7}
                   (\M{1}{0}) node[right] {$\alpha_{2}$};

      \draw[beta]  (\C{2}) node[above left] {7} -| (\S{2}{0}) node[below right] {7}
                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {7\,}
                   (\M{2}{0}) node[right] {$\beta_7$};
      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {15} -| (\S{3}{0}) node[above right] {\!20}
                   (\M{2}{1}) node[right] {$\alpha_{15,1,27}$};
      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {1} -| ([xshift=-1pt]\S{3}{0}) node[above left] {6\!};
      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,27} -| ([xshift=-3pt]\S{3}{0}) node[below] {0\,\,};
      \draw[gamma] (\S{2}{1}) node[above] {\,7} -- (\S{3}{1}) node[above left] {6\!}
                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above] {7\,}
                   (\M{2}{2}) node[right] {$\gamma_7$};
      \draw[delta] (\S{2}{4}) node[above] {\,7} -- (\S{3}{4}) node[above right] {20}
                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below] {\,7}
                   (\M{2}{4}) node[right] {$\delta_7$};
      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {7\,}
                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-4pt]\S{3}{0}) node[above left] {0}
                   (\M{2}{3}) node[right] {$\varepsilon_0$};

      \draw[beta]  (\C{3}) node[above left] {6} -| ([xshift=-1pt]\S{3}{0}) node[below left] {6\!}
                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {6\!}
                   (\M{3}{0}) node[right] {$\beta_{20,6}$};
      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {14} -| (\S{4}{0}) node[above right] {19}
                   (\M{3}{1}) node[right] {$\alpha_{14}$};
      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {20} -- ([xshift=2pt]\S{4}{1}) node[above right] {19}
                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {20}
                   (\M{3}{2}) node[right] {$\gamma_{20}$};

      \draw[beta]  (\C{4}) node[left] {19} -| (\S{4}{0}) node[below right] {19}
                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {19}
                   (\M{4}{0}) node[right] {$\beta_{19}$};

      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
      \draw (\S{0}{2}) node[below] {\normalsize \MiniMORUS[640]};
    \end{scope}

    \draw[dashed] (\S{4.25}{-1}) -- (\S{4.25}{7}) -- ++(\S{1}{-1}) -- ++(\S{0}{7});

    \begin{scope}[yshift=-7.0cm]
      \foreach \r in {0,...,3} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
      \foreach \r in {0,...,4} { \draw (\S{0}{\r}) node[above] {$S_{\r}$}; }
      \draw (\C{0}|-\S{0}{0}) node[above] {$C$};

      \draw[alpha] (\C{0}) node[left] {51} -| (\S{1}{0}) node[above right] {0}
                   (\M{0}{0}) node[right] {$\alpha_{51}$};

      \draw[beta]  (\C{1}) node[above left] {0} -| (\S{1}{0}) node[below right] {0}
                   (\C{1}-|\S{1}{0}) -| ([xshift=-1pt]\S{1}{1}) node[above] {0\,}
                   (\M{1}{0}) node[right] {$\beta_0$};
      \draw[alpha] ([yshift=-1.5pt]\C{1}) node[left] {55} -| (\S{2}{0}) node[above right] {\!4}
                   (\M{1}{1}) node[right] {$\alpha_{55,33}$};
      \draw[alpha] ([yshift=-3.0pt]\C{1}) node[below left] {33} -| ([xshift=-1pt]\S{2}{0}) node[above left] {46\!};
      \draw[gamma] (\S{1}{1}) node[above] {\,0} -- (\S{2}{1}) node[above left] {46\!}
                   (\S{1}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{1}{4}) node[above] {0\,}
                   (\M{1}{2}) node[right] {$\gamma_0$};
      \draw[delta] (\S{1}{4}) node[above] {\,0} -- (\S{2}{4}) node[above right] {4}
                   (\S{1}{4}) ++(0,-1.25) -| ([xshift=2pt]\S{2}{2}) node[below] {0} node[black] {$\times$}
                   (\M{1}{3}) node[right] {$\delta_0$};

      \draw[beta]  (\C{2}) node[above left] {46} -| ([xshift=-1pt]\S{2}{0}) node[below left] {46\!}
                   ([xshift=-1pt]\C{2}-|\S{2}{0}) -| (\S{2}{1}) node[below left] {46\!}
                   (\M{2}{0}) node[right] {$\beta_{4,46}$};
      \draw[beta]  ([yshift=-1.5pt]\C{2}) node[left] {4} -| (\S{2}{0}) node[below right] {\!4}
                   ([yshift=-1.5pt]\C{2}-|\S{2}{0}) -| ([xshift=1pt]\S{2}{1}) node[above] {\,\,\,4};
      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {37} -| (\S{3}{0}) node[above right] {50}
                   (\M{2}{1}) node[right] {$\alpha_{37}$};
      \draw[gamma] ([xshift=2pt]\S{2}{1}) node[below right] {4} -- ([xshift=2pt]\S{3}{1}) node[above right] {50}
                   ([xshift=2pt]\S{2}{1}) ++(0,-.75) -| (\S{2}{4}) node[below right] {4}
                   (\M{2}{2}) node[right] {$\gamma_{4}$};

      \draw[beta]  (\C{3}) node[left] {50} -| (\S{3}{0}) node[below right] {50}
                   (\C{3}-|\S{3}{0}) -| ([xshift=2pt]\S{3}{1}) node[below right] {50}
                   (\M{3}{0}) node[right] {$\beta_{50}$};

      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
      \draw (\S{5}{2}) node[above] {\normalsize \MiniMORUS[1280]};
    \end{scope}

    \begin{scope}[xshift=2.5cm,yshift=-7.0cm]
      \foreach \r in {1,...,4} { \draw[gray, rounded corners=2pt] (0,-\r*1.5) rectangle ++(1.5,-1.5); }
      \foreach \r in {0,...,4} { \draw (\S{1}{\r}) node[above] {$S_{\r}$}; }
      \draw (\C{0}|-\S{1}{0}) node[above] {$C$};

      \draw[alpha] (\C{1}) node[left] {25} -| (\S{2}{0}) node[above right] {38}
                   (\M{1}{0}) node[right] {$\alpha_{25}$};

      \draw[beta]  (\C{2}) node[above left] {38} -| (\S{2}{0}) node[below right] {38}
                   (\C{2}-|\S{2}{0}) -| ([xshift=-1pt]\S{2}{1}) node[above] {38\,}
                   (\M{2}{0}) node[right] {$\beta_{38}$};
      \draw[alpha] ([yshift=-1.5pt]\C{2}) node[left] {7} -| (\S{3}{0}) node[above right] {\!20}
                   (\M{2}{1}) node[right] {$\alpha_{7,1,51}$};
      \draw[alpha] ([yshift=-3.0pt]\C{2}) node[below left] {29} -| ([xshift=-1pt]\S{3}{0}) node[above left] {42\!};
      \draw[alpha] ([yshift=-4.5pt]\C{2}) node[below] {\,\,\,\,51} -| ([xshift=-4pt]\S{3}{0}) node[below] {0\,\,\,};
      \draw[gamma] (\S{2}{1}) node[above right] {38} -- (\S{3}{1}) node[above left] {20\!}
                   (\S{2}{1}) ++(0,-.75) -| ([xshift=-1pt]\S{2}{4}) node[above left] {38}
                   (\M{2}{2}) node[right] {$\gamma_{38}$};
      \draw[delta] (\S{2}{4}) node[above] {\,38} -- (\S{3}{4}) node[above right] {42}
                   (\S{2}{4}) ++(0,-1.25) -| ([xshift=1pt]\S{3}{2}) node[below right] {38}
                   (\M{2}{4}) node[right] {$\delta_{38}$};
      \draw[epsil] (\S{2}{2}) node[above] {0} node[black] {$\times$} -- (\S{3}{2}) node[below] {38}
                   (\S{2}{2}) ++(0,-1.00) -| ([xshift=-5pt]\S{3}{0}) node[above left] {0}
                   (\M{2}{3}) node[right] {$\varepsilon_0$};

      \draw[beta]  (\C{3}) node[above left] {42} -| ([xshift=-1pt]\S{3}{0}) node[below left] {42\!}
                   ([xshift=-1pt]\C{3}-|\S{3}{0}) -| (\S{3}{1}) node[below left] {42\!}
                   (\M{3}{0}) node[right] {$\beta_{20,42}$};
      \draw[beta]  ([yshift=-1.5pt]\C{3}) node[left] {20} -| (\S{3}{0}) node[below right] {\!20}
                   ([yshift=-1.5pt]\C{3}-|\S{3}{0}) -| ([xshift=1pt]\S{3}{1}) node[above] {\,\,\,\,20};
      \draw[alpha] ([yshift=-3.0pt]\C{3}) node[below left] {11} -| (\S{4}{0}) node[above right] {24}
                   (\M{3}{1}) node[right] {$\alpha_{11}$};
      \draw[gamma] ([xshift=2pt]\S{3}{1}) node[below right] {42} -- ([xshift=2pt]\S{4}{1}) node[above right] {24}
                   ([xshift=2pt]\S{3}{1}) ++(0,-.75) -| (\S{3}{4}) node[below right] {42}
                   (\M{3}{2}) node[right] {$\gamma_{42}$};

      \draw[beta]  (\C{4}) node[left] {24} -| (\S{4}{0}) node[below right] {24}
                   (\C{4}-|\S{4}{0}) -| ([xshift=2pt]\S{4}{1}) node[below right] {24}
                   (\M{4}{0}) node[right] {$\beta_{24}$};

      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
    \end{scope}
  \end{tikzpicture}}

  \caption{\MiniMORUS: two approximations for $S_{2,0}^2$.}
  \label{fig:twoapproximations}
\end{figure}
